---
title: "360buy Case Study Report"
output: html_document
---
# Necessary libraries
```{r}
library(readxl)
library(dplyr)
library(factoextra)
library(ggplot2)
library(cluster)
library(gt)
library(tidyr)
```
# Reading the Data
```{r}
data <- read_excel("~/Downloads/360buy_SurveyData.xlsx")
head(data) 
summary(data)
str(data)
```
# Selecting the Segmentation Variables
```{r}
data_segmentation <- data %>%
  select(CusChoice, ConstUp, ReplacReminder, ProInsuCov, ProdReturn)
head(data_segmentation)
summary(data_segmentation)
str(data_segmentation)
```
# Selecting the Demographic Variables
```{r}
data_demographic <- data %>%
  select(CusAgeYr, CusGen, LevEdn, LevIncome, CusAcct)
head(data_demographic)
summary(data_demographic)
str(data_demographic)
```
#Scaling the Data
```{r}
data_scaled_seg <- scale(data_segmentation)
apply(data_scaled_seg, 2, mean)
apply(data_scaled_seg, 2, sd)
head(data_scaled_seg)
```
# Scree Plot
```{r}
fviz_nbclust(data_scaled_seg, kmeans, method = 'wss')
```
# Heirachical Clustering
```{r}
dist_data_seg <- dist(data_scaled_seg, method = 'euclidean')
hc_data_seg <- hclust(dist_data_seg, method ='ward.D2')
cluster <- cutree(hc_data_seg, k = 3)

plot(hc_data_seg, cex = 0.5, lwd = 0.5, main = 'Cluster Dendogram', xlab = 'Respondents', sub = '')
rect.hclust(hc_data_seg, k=3)
```
# Visualisation of Clusters
```{r}
fviz_cluster(list(data = data_scaled_seg, cluster = cluster))
```
# Binding the Cluster to Whole Data
```{r}
data_2 <- data
data_2$cluster <- cluster
table(data_2$cluster)
```
# Binding the Cluster to Scaled Segmented Data
```{r}
data_scaled_seg <- as.data.frame(data_scaled_seg)
data_scaled_seg$cluster <- cluster
table(data_scaled_seg$cluster)
```
# Binding the Cluster to Unscaled Segmented Data
```{r}
data_segmentation <- as.data.frame(data_segmentation)
data_segmentation$cluster <- cluster
table(data_segmentation$cluster)
```
# Binding the Cluster to Demographic Data
```{r}
data_demographic <- as.data.frame(data_demographic)
data_demographic$cluster <- cluster
table(data_demographic$cluster)
```
# Finding the Proportions of the Clusters
```{r}
proportions_seg <- table(data_2$cluster) / length(data_2$cluster) * 100
proportions_seg
```
# Mean and Standard Deviation of Segments
```{r}
segments <- data_segmentation %>% group_by(cluster)%>%summarise(across(everything(), mean))
mean_table <- segments %>% gt() %>% tab_header (title = md('Mean Values for Clusters'))
mean_table
```
```{r}
segments_2 <- data_segmentation %>% group_by(cluster)%>%summarise(across(everything(), median))
median_table <- segments_2 %>% gt() %>% tab_header (title = md('Median Values for Clusters'))
median_table
```
```{r}
segments_3 <- data_segmentation %>% group_by(cluster)%>%summarise(across(everything(), sd))
sd_table <- segments_3 %>% gt() %>% tab_header (title = md('Standard Deviation Values for Clusters'))
sd_table
```
# Comparing Demographic Variables across Clusters
```{r}
segments_d <- data_demographic %>% group_by(cluster) %>% summarise(across(everything(), mean))
mean_table_d <- segments_d %>% gt() %>% tab_header (title = md('Mean Demographics across Clusters'))
mean_table_d
```
```{r}
segments_d2 <- data_demographic %>% group_by(cluster)%>%summarise(across(everything(), median))
median_table_d <- segments_d2 %>% gt() %>% tab_header (title = md('Median Demographics across Clusters'))
median_table_d
```
```{r}
segments_d3 <- data_demographic %>% group_by(cluster)%>%summarise(across(everything(), sd))
sd_table_d <- segments_d3 %>% gt() %>% tab_header (title = md('Standard Deviation of Demographic Variables across Clusters'))
sd_table_d
```
# Cluster Names - Demographic - For Barchart
```{r}
cluster_names <- c("Price-sensitive Consumers", "Affluent-Upgraders", "Digitally Engaged Responders")
data_2$cluster_label <- factor(data_2$cluster, levels = c(1, 2, 3), labels = cluster_names)
```
# Bar Chart - Mean Cluster Characteristics with Standard Deviations
```{r}
data_segmentation$cluster_label <- data_2$cluster_label

data_long <- data_segmentation %>% pivot_longer(cols = c(CusChoice, ConstUp, ReplacReminder, ProdReturn, ProInsuCov), names_to = "variable", values_to = "value")

summary_stats <- data_long %>% group_by(cluster_label, variable) %>% summarise(mean = mean(value), sd = sd(value), .groups = "drop")

ggplot(summary_stats, aes(x = variable, y = mean, fill = cluster_label)) + geom_bar(stat = "identity", position = "dodge") + geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), position = position_dodge(0.9), width = 0.2) + labs(title = 'Behavioural Differences across Clusters (Mean + SD)', x = 'Behavioural Variables', y = 'Mean Value', fill = 'Cluster') + theme_minimal()
```
# Box Plot - Demographic Variables 
```{r}
data_demographic$cluster_label <- data_2$cluster_label

data_long <- data_demographic %>% pivot_longer(cols = c(CusAgeYr, LevEdn, LevIncome), names_to = "variable", values_to = "value")

ggplot(data_long, aes(x = cluster_label, y = value, fill = variable)) + geom_boxplot(position = position_dodge(width = 0.8)) + labs(title = "Demographic Variables by Cluster", x = "Cluster", y = "Value", fill = "Variable") + theme_minimal()
```
# Bar Chart - Demographic Variables - Categorical
```{r}
data_long_2 <- data_demographic %>% pivot_longer(cols = c(CusGen, CusAcct), names_to = "variable",values_to = "value")

summary_stats_2 <- data_long_2 %>% group_by(cluster_label, variable) %>% summarise(mean = mean(value), .groups = "drop")

ggplot(summary_stats_2, aes(x = variable, y = mean, fill = cluster_label)) + geom_bar(stat = "identity", position = "dodge") + labs(title = "Demographic Variables by Cluster",x = "Demographic Variable",y = "Proportion",fill = "Cluster") + theme_minimal()
```
# Pie Chart - Size of Each Cluster
```{r}
cluster_sizes <- as.data.frame(proportions_seg)
names(cluster_sizes) <- c("cluster_label", "proportion")

ggplot(cluster_sizes, aes(x = "", y = proportion, fill = cluster_label)) + geom_bar(stat = "identity", width = 1) + coord_polar(theta = "y") + theme_void() + geom_text(aes(label = paste0(round(proportion, 1), "%")), position = position_stack(vjust = 0.5)) +labs(title = "Relative Size of Market Clusters", fill = "")
```



